---
title: "HotelsFinalAnalysis"
author: "David Snider"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(countrycode)
library(rvest)
library(dplyr)
library(tidyr)
library(xtable)
library(modelr)
library(broom)
library(class)
library(lubridate)
library(pracma)
library(class)

df = read_csv("hotel_booking.csv")
```

# Question 2: Predicting  Cancellation w/ KNN

Formulas
```{r}
standardize = function(vector) {
  z = (vector-mean(vector))/sd(vector)
  return(z)
}

#divides correct prediction by total predictions
accuracy = function(x) {
  sum(diag(x)/ sum(diag(x)/(sum(rowSums(x))))) *100
}
```

```{r}
tab = matrix(seq(1,100), ncol = 1, byrow = F)
colnames(tab) = c("Index")
```


Randomly obtain small subset of data:
```{r}
set.seed(216)
ran = sample(1:nrow(df), 0.9 * nrow(df))

df = df %>% mutate(hotel = ifelse(hotel == "Resort Hotel", 1, 0))
df_norm = as.data.frame(lapply(df[, c("previous_cancellations", "previous_bookings_not_canceled", "lead_time", "adr")], standardize)) #including hotel causes NANs

#subset = df[sample(x=nrow(df), size=10000), ]

df_train = df_norm[ran, ] #%>% cbind(., )
df_test = df_norm[-ran, ]

df_target_category = df[ran, "is_canceled"]
df_test_category = df[-ran, "is_canceled"]
```

```{r}
df_train %>% is.na()
```

```{r}
acc = vector("list", 100)

for (k in 1:100){
  knn_output = knn(df_train, df_test, cl=df_target_category, k=k, prob=T)
  confusion_matrix = table(knn_output, df_test_category)
  
  acc[[k]]= accuracy(confusion_matrix)
  
}
```


```{r}
tab = cbind(tab, acc)
tab
```


Use cross validation to find accuracy for each k-value on that subset (maybe a range of k from 1 to 100)
```{r}

df_knn = subset %>%
  mutate( #modified test set to standardize all variables
    previous_cancellations= standardize(previous_cancellations),
    previous_bookings_not_canceled= standardize(previous_bookings_not_canceled),
    hotel= ifelse(hotel=="Resort Hotel", 1, 0), #still have to do
    hotel= standardize(hotel),
    lead_time= standardize(lead_time),
    adr= standardize(adr)) %>% 
  mutate(predicted_canc4= knn(
    train= dplyr::select(df_train, lead_time, adr, previous_cancellations),
    test= dplyr::select(df_test, lead_time, adr, previous_cancellations),
    cl= factor(df_train$is_canceled, levels=c(0,1), labels=c("Not canceled", "Canceled")
        ), 
    k=5)) %>% 
  dplyr::select(predicted_canc4)

#function to give 
model.knn.func = function(data, k) {
  
}

for (k in 1:50) {
  for (c in 1:5) {
    
  }
  #accuracy of k = mean(accuracy for each c)
}
```
Steps:

Make sure there aren't too many NA's (check)

Get code for cross-validated accuracy
```{r}
subset.c= subset %>% crossv_kfold(k=5)
subset1 = subset.c[1,]
subset1 %>% select(train) %>%
  unnest(cols=c(train))

#cross validation code
for (c in 1:5) {
  #make predictions using knn
  df_test %>% 
  mutate( #modified test set to standardize all variables
    previous_cancellations=standardize(previous_cancellations),
    previous_bookings_not_canceled=standardize(previous_bookings_not_canceled),
    hotel=ifelse(hotel=="Resort Hotel", 1, 0),
    hotel=standardize(hotel),
    lead_time=standardize(lead_time),
    adr=standardize(adr)
    ) %>% 
  mutate(prediction=knn(
    train=dplyr::select(
      df_train,
      lead_time,
      adr,
      previous_cancellations
      ),
    test=dplyr::select(
      df_test,
      lead_time,
      adr,
      previous_cancellations
    ),
    cl=factor(
      df_train$is_canceled,
      levels=c(0,1),
      labels=c("Not canceled","Canceled")
    ),
    k=2
  )
  
  #calculate accuracy using predictions
}
```


Identify best 3 k-values
```{r}

```

Run those best 3 k-values on the entire data (can split into train-test, i.e. no need for cross validation)
```{r}

```


Identify best k-value
```{r}

```

Weighting best k-NN with simple log to get desired specificity/sensitivity
```{r}

```


