---
title: "HotelsFinalAnalysis"
author: "David Snider"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(countrycode)
library(rvest)
library(dplyr)
library(tidyr)
library(xtable)
library(modelr)
library(broom)
library(class)
library(lubridate)
library(pracma)

df = read_csv("hotel_booking.csv")
```


# Question 2:

Randomly obtain small subset of data:
```{r}
set.seed(216)
subset = df[sample(x=nrow(df), size=10000),]
```

Select a small subset of observations from the dataset (so that we can run multiple k-NN's without the computer crashing) 
```{r}

```


Use cross validation to find accuracy for each k-value on that subset (maybe a range of k from 1 to 100)
```{r}

```


Identify best 3 k-values
```{r}

```

Run those best 3 k-values on the entire data (can split into train-test, i.e. no need for cross validation)
```{r}

```


Identify best k-value
```{r}

```

Weighting best k-NN with simple log to get desired specificity/sensitivity
```{r}

```


