---
title: "HotelsFinalAnalysis"
author: "David Snider"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(countrycode)
library(rvest)
library(dplyr)
library(tidyr)
library(xtable)
library(modelr)
library(broom)
library(class)
library(lubridate)
library(pracma)
library(class)

df = read_csv("hotel_booking.csv")
```

# Question 2: Predicting  Cancellation w/ KNN

Formulas
```{r}
standardize = function(vector) {
  z = (vector-mean(vector))/sd(vector)
  return(z)
}

#divides correct prediction by total predictions
accuracy = function(x) {
  sum(diag(x)/(sum(rowSums(x)))) * 100
}


```

```{r}
tab = matrix(seq(1,100), ncol = 1, byrow = F)
colnames(tab) = c("Index")
```


Randomly obtain small subset of data:
```{r}
set.seed(216)
ran = sample(1:nrow(df), 0.9 * nrow(df))

df = df %>% mutate(hotel = ifelse(hotel == "Resort Hotel", 1, 0))
df_norm = as.data.frame(lapply(df[, c("previous_cancellations", "previous_bookings_not_canceled", "lead_time", "adr")], standardize)) #including hotel causes NANs

#subset = df[sample(x=nrow(df), size=10000), ]

df_train = df_norm[ran, ] #%>% cbind(., )
df_test = df_norm[-ran, ]

df_target_category = df[ran, "is_canceled"] %>% pull(is_canceled)
df_test_category = df[-ran, "is_canceled"] %>% pull(is_canceled)
```

```{r}
acc = list()
options(width = 80)

for (k in 1:20){
  knn_output = knn(df_train, df_test, cl=df_target_category, k=k)
  confusion_matrix = table(knn_output, df_test_category)
  
  acc[[k]]= accuracy(confusion_matrix)
}
```


```{r}
tab = cbind(tab, acc)
tab %>% unnest() %>% as.data.frame()
```

% of people that cancelled their hotel booking
```{r}
df$is_canceled %>% sum()/ count(df)
```


```{r}
ggplot(as.data.frame(tab)) + geom_line(aes(x=Index, y=acc))
```


Use cross validation to find accuracy for each k-value on that subset (maybe a range of k from 1 to 100)

# Next Steps:

- Visualize Metrics as K increases for KNN
- Calculate F1 score for both Logistic Regression and KNN + Visualize
- Make a RF model w/ accomopanying metrics

Make sure there aren't too many NA's (check)

