---
title: "HotelsFinalAnalysis"
author: "David Snider"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(countrycode)
library(rvest)
library(dplyr)
library(tidyr)
library(xtable)
library(modelr)
library(broom)
library(class)
library(lubridate)
library(pracma)
library(class)
df = read_csv("hotel_booking.csv")
k_df = read_csv("k_df.csv")
```

# Question 2: Predicting  Cancellation w/ KNN

Formulas
```{r}
standardize = function(vector) {
  z = (vector-mean(vector))/sd(vector)
  return(z)
}
#divides correct prediction by total predictions
accuracy = function(x) {
  sum(diag(x)/(sum(rowSums(x))))
}
```

Created file, so no longer need to run this code. Storing it here with eval=F.
```{r, eval=F}
set.seed(216)
ran = sample(1:nrow(df), 0.8 * nrow(df))
df_hotel_binary = df %>% mutate(hotel = ifelse(hotel == "Resort Hotel", 1, 0))
df_norm = as.data.frame(lapply(df_hotel_binary[, c("previous_cancellations", "previous_bookings_not_canceled", "lead_time", "adr", "hotel")], standardize)) 
#subset = df[sample(x=nrow(df), size=10000), ]
df_train = df_norm[ran, ] #%>% cbind(., )
df_test = df_norm[-ran, ]
df_target_category = df[ran, "is_canceled"] %>% pull(is_canceled)
df_test_category = df[-ran, "is_canceled"] %>% pull(is_canceled)
acc = vector("double", 100)
sens = vector("double", 100)
spec = vector("double", 100)
options(width = 80)
for (k in 1:100){
  knn_output = knn(df_train, df_test, cl=df_target_category, k=k)
  confusion_matrix = table(knn_output, df_test_category)
  acc[[k]]= accuracy(confusion_matrix)
  sens[[k]]= confusion_matrix[2,2] / sum(confusion_matrix[,2])
  spec[[k]]= confusion_matrix[1,1] / sum(confusion_matrix[,1])
}
k = as.integer(1:100)
k_df = data.frame(
  k,
  acc,
  sens,
  spec
)
```

% of people that cancelled their hotel booking
```{r}
sum(df$is_canceled)/ count(df)
```

Use cross validation to find accuracy for each k-value on that subset (maybe a range of k from 1 to 100)

# Next Steps:

- Visualize Metrics as K increases for KNN
```{r}
k_df_plot = k_df %>% 
  rename(
    "Accuracy"="acc",
    "Sensitivity"="sens",
    "Specificity"="spec"
  ) %>%
  gather(
    "Accuracy":"Specificity", 
    key="Metric", 
    value="Value"
    )
ggplot(k_df_plot) + 
  geom_line(aes(x=k, y=Value, color=Metric)) + 
  ggtitle("Performance of k-NN model by k value")
```

- Calculate F1 score for both Logistic Regression and KNN + Visualize
- Make a RF model w/ accomopanying metrics

Make sure there aren't too many NA's (check)
```{r}
city_hotels2 = df %>% filter(hotel == "City Hotel", is_canceled == 0) %>% dplyr::select(arrival_date_year, arrival_date_month, arrival_date_day_of_month,stays_in_week_nights, stays_in_weekend_nights, adr) %>% mutate(length_of_stay = stays_in_week_nights + stays_in_weekend_nights) %>% arrange(arrival_date_year, arrival_date_month, arrival_date_day_of_month)

city_hotels2$arrival_date_month = as.integer(factor(city_hotels2$arrival_date_month, levels = month.name))

city_hotels2 = city_hotels2 %>% dplyr::select(arrival_date_year, arrival_date_month, arrival_date_day_of_month, length_of_stay, adr) %>% unite("arrival_date", c("arrival_date_year", "arrival_date_month", "arrival_date_day_of_month"), sep = "/") %>% filter(adr != 0.00)

city_hotels2$arrival_date <- as.Date(city_hotels2$arrival_date)
city_hotels2 = city_hotels2 %>% arrange(arrival_date)  %>% group_by(arrival_date) %>% summarise(avg_length_of_stay = mean(length_of_stay), avg_adr = mean(adr))

model_data_city = city_hotels2
start_date = as.Date("2015-07-01")
model_data_city$arrival_date <- as.numeric(difftime(model_data_city$arrival_date, start_date, unit = "days"))
NumDays.city <- model_data_city$arrival_date
```

```{r, results="asis"}
q1table = model_data_city %>% 
  head(4) %>%
  select(-avg_length_of_stay) %>%
  mutate(arrival_date = as.integer(arrival_date)) %>%
  xtable(
    align="ccc", 
    type="html")
print(q1table, 
      "html", 
      include.rownames=FALSE,
      html.table.attributes="align='center',
                             rules='rows',
                             width=50%,
                             frame='hsides',
                             border-spacing=5px")
```
